Landau quantization of nearly degenerate bands

### 2018-10-02 paper.tex
Updates from ShareLaTeX

\documentclass[aps, prb, showpacs, twocolumn, notitlepage, superscriptaddress]{revtex4-1}
\usepackage{amsmath}
\usepackage{amssymb,amscd,bm,dsfont,wasysym,mathrsfs,latexsym,psfrag,accents,mathtools}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{color}
\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{boldline}
\usepackage[normalem]{ulem}
\input{preamble}

@@

\begin{abstract}
\textbf{[abstract needs revamp]}
 In the semiclassical theory of Landau levels, a Bloch electron is driven by a magnetic field  along constant-energy band contours (in quasimomentum space) known as orbits. We introduce a quantization rule for Landau levels of nearly degenerate orbits which are split by spin-orbit coupling. The average area $S$ of two spin-orbit-split orbits is assumed much larger than the difference in areas ($\delta S$), and also much larger than the inverse square of the magnetic length ($\lmt$). By treating both $\delta S/S$ and $1/l^2S$ as small parameters with finite ratio $l^2\delta S$, our perspective diverges from and generalizes previous semiclassical theories.  $l^2\delta S$ is the dimensionless measure of the field-induced hybridization between spin-orbit-split orbits; we allow for $l^2\delta S$ to be comparable with the Zeeman coupling. In the opposing regimes $l^2\delta S {{\ll}} 1$  and $l^2\delta S{\gg}1$, we recover the known Onsager-Lifshitz-Roth rules for nondegenerate and degenerate orbits, respectively.  In the absence of crystallographic symmetry, three real parameters are needed to tune a degeneracy between two Landau levels which have been split by the spin-orbit and Zeeman interactions. Here, we identify symmetry classes for which such degeneracies can be tuned by varying a single parameter, which might be the magnitude or orientation of the field, or a bias voltage in tunneling spectroscopy.  We demonstrate the utility of our rule for the Rashba 2DEG (for which an in-plane Zeeman field can induce a type-II Dirac point), Dresselhaus spin-orbit-coupled systems, and the angular magnetoresistance oscillations of semiconductor bilayer heterostructures.      
\end{abstract}
\date{\today}

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RZ.png}
\caption{(a) Trefoille-shaped cyclotron orbit for the Rashba-Zeeman model at the energy of the Dirac point. The orbit is drawn over the  spin-splitting field ($d_x,d_y$) of the degeneracy-lifting Hamiltonian $\delta\var$ [cf.\ \q{eq:RZ-Hamiltonian}]. The wavefunction along the inner (resp.\ outer) component of the trefoille orbit has its pseudospin aligned parallel (resp.\ anti-parallel) to the spin-splitting field. (b) Upper panel: a schematic illustration of Landau levels near the II-Dirac point of the Rashba model (with in-plane Zeeman field, and in the weak- hybridization limit). The solid line represents the Onsager-Lifshitz quantization of the trefoille orbit [cf.\ Eq.\ ({trefoillerule})], and the dashed line incorporates the first-order correction [cf.\ Eq.\ (\ref{RZfirstorderE})] due to tunneling. 
Lower panel: oscillation of $\delta E_n$ [cf. Eq. (\ref{RZfirstorderE})] for even integer $n$ with respect to magnetic field calculated by Eq. (\ref{RZfirstorderE}) (black dashed line) and numerical diagonalization (red solid line). $\delta E$ is extracted at the Landau level closest to 0.45eV and the II-Dirac point is located at 0.5eV. Parameters for lower panel of (b): $m{=}0.076m_0$, $\alpha{=} 1.5(10^{5}$cm/s$\cdot\hbar$), $Z_\parallel {=}$1meV. Out-of-plane Zeeman coupling is negligibly weak for the choice of $m$. (c) A sketch of cyclotron orbits of Rashba Hamiltonian with in-plane magnetic field. Inset: an enlarged view of near II-Dirac point. (d) For various configurations ($\alpha$, $Z_\parallel$) of $H_{RZ}$, we plot in left panel $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that is closest to $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}), where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian $H_{RZ}(\bK)$. Corresponding $l^2 \delta S^2/S$ are plotted in right panel. We only plotted in region where $0.4{<}\epsilon_0{<}1.0$(eV).\cite{RZfignote}, Units: $\alpha$($1.5\times 10^{5}$cm/s$\cdot\hbar$). Parameters for (d): $m=0.076m_0$, $l=100$\AA. \label{fig:RZ}}.
\end{figure}

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{blochsphere.png}
\caption{Solid-angle representation of  the single-band Berry phase $\phi_{\pm}^B$ for inner (subscript $-$, upper panel) and outer ($+$, lower panel) orbit, at various energies close to the Dirac point $\epsilon_0$ [cf.\ Eq. (\ref{whereisdiracpoint})]. The decomposition of the effective Hamiltonian [cf.\ Eq. (\ref{calHRZ})] into  Pauli matrices ($\calh{=}\sum_j\calh_j\sigma_j$) defines a three-vector $\boldsymbol{\calh}{=}({-}\alpha k_y$,$Z_\parallel{+}\alpha k_x,{-}Z_\perp$). $\phi_{\pm}^B$ equals half the solid angle subtended by   ${\mp} \boldsymbol{\calh}(\bk)$ [solid blue line] as $\bk$ is varied along the outer/inner orbit;[Berry] the orientation of the solid angle is indicated by blue dotted lines. The out-of-plane Zeeman interaction lifts $\boldsymbol{\calh}$ out of the x-y plane, but this occurs at most over a fraction ${\sim}(\bar{\mu}_{\sma{\text{min}}}m_0/m{\ll}1)$ of the cyclotron period. \label{fig:blochsphere}}
\end{figure}

@@

 To generalize the above discussion, let us parametrize ${\cal \bar{A}}$ by $d$ independent real numbers  $\{r_1,\ldots,r_{d}\}{:}{=}\br $;  ${\cal \bar{A}}(\br){=}{\pm}I$ defines a set of submanifolds in $\br$-space which we refer to as degeneracy manifolds; in the above example, these manifolds comprise two points on $S^3$. The codimension of the degeneracy manifold  is defined as $d$ minus the dimension of said manifold; alternatively stated, the codimension is the number of linearly-independent parameters needed to fine-tune a degeneracy. Here and henceforth, we adopt the shorthand that  \textit{parameters} will always refer to real parameters, and \textit{codimension} always refers to a codimension of a degeneracy manifold.

@@

At energies close to the Dirac point ($E{-}\epsilon_0 {\ll} wm_2(2\epsilon_0/m_1)^{\sma{1/2}}$), there are two concentric orbits which (nearly) touch at both Dirac points, as illustrated in Fig. \ref{fig:Cn-breakdown} (a). For sufficiently weak field, interband tunneling is localized near either II-Dirac point, and occurs with the Landau-Zener probability:
\e{\rho^2=e^{-2\pi\bar{\mu}}, \; \bar{\mu}=\f{\big(Z_{\perp}^2+(m_1/m_2)^2(E-\epsilon_0)^2\big)l^2}{2w\sqrt{2E/m_1}}. \label{landauzenerc2}}
Equivalently, $\bar{\mu}{=}S_{\sma{\square}}l^2/8$, with $S_{\sma{\square}}$ the area of rectangle inscribed between the two hyperbolic arms [cf.\  Fig\ref{fig:Cn-breakdown}(a)]. Elsewhere, an electron undergoes adiabatic (band-conserving) dynamics. Due to $\mathfrak{c}_2$ symmetry, the time-evolution propagator over the full cyclotron period is simply the square of the propagator for half the period: 
\e{\A=\left(\matrixtwo{\tau e^{i\bar{\varphi}}}{-\rho}{\rho}{\tau e^{-i\bar{\varphi}}}{\diagmatrix{e^{i\tilde{\lambda}_-}}{e^{i\tilde{\lambda}_+}}} \right)^2. \label{propC2breakdown}}
The half-period propagator may be taken from \qq{scattmat}{propinplanezeeman}, with  $\bar{\mu}$ here given by \q{landauzenerc2}, and  $\tilde{\lambda}_{\pm}{=}{\pm}l^2\delta S/4$ plus a field-independent geometric phase.

@@

\begin{table*}
\begin{tabular*}{\textwidth}{c@{\extracolsep{\fill}}ccccccc}
\hlineB{2.0}
             class number & action on orbit & action on $\bk$ & $u$ & $s$ & quasidegenerate & $g$ & codimension\\
\hline
\multirow{2}{*}{1} & \multirow{10}{*}{$|g\circ \frako_0| = |\frako_0|$} & \multirow{4}{*}{$\forall \bk \in \frako_0$, $g\circ \bk=\bk$} & \multirow{2}{*}{0} & \multirow{2}{*}{0} & \multirow{2}{*}{$\surd$} & $\breve{g}\propto I$ & 3  \\
 & & & & & & $\breve{g}\not\propto I$ & 1 \\
 \cline{4-8}
\multirow{2}{*}{2} & & & \multirow{2}{*}{0} & \multirow{2}{*}{1} & & $g^2=1$ & 1  \\
 & & &  &  & & $g^2=-1$ & 3  \\
 \cline{3-8}
3 & & \multirow{6}{*}{$\exists \bk \in \frako_0, g\circ \bk\ne\bk$} & 0 & 0 & $\surd$ & - & 1   \\
\cline{4-8}
4 & & & 0 & 1 & & - & 1   \\
\cline{4-8}
\multirow{2}{*}{5} & & & \multirow{2}{*}{1} & \multirow{2}{*}{0} & & $\breve{g} \propto \sigma_z$ & 2  \\
 & & & &  & & $\breve{g} \not\propto \sigma_z$ & 0  \\
\cline{4-8}
\multirow{2}{*}{6} & & & \multirow{2}{*}{1} & \multirow{2}{*}{1} & \multirow{2}{*}{$\surd$} & $g^2=1$ & 2  \\
& & & &  & & $g^2=-1$ & 0  \\
\hline
7 & \multirow{4}{*}{$|g\circ \frako_0| \ne |\frako_0|$} & \multirow{4}{*}{-} & 0 & 0 & $\surd$ & - & 3 \\
\cline{4-8}
8 &  &  & 0 & 1 & & - & 3 \\
\cline{4-8}
9 &  &  & 1 & 0 & & - & 3 \\
\cline{4-8}
10 &  &  & 1 & 1 & $\surd$ & - & 3 \\
\hlineB{2.0}
\end{tabular*}
\caption{Symmetry-constrained codimensions for the two-by-two propagator $\A$. The first column numbers the symmetry class of $\A$. The second and third column defines how the symmetry acts on the cyclotron orbit. The next two columns  inform us if the mapping $\bk{\rightarrow}g{\circ}\bk$ [cf.\ \q{gmapk}]  preserves ($u{=}0$) or inverts ($u{=}0$) the orientation of the orbit, and if $g$ preserves ($s{=}0$) or inverts ($s{=}1$) the arrow of time. The sixth column indicates whether this  class is potentially a symmetry of quasidegenerate orbits. The seventh column specifies additional conditions (on the representation of $g$) which must be imposed to uniquely determine the codimension. $\breve{g}$ is defined as the diagonalized, unitary matrix representation of $g$ in the degenerate subspace of $\hat{H}_0$. In class 5, $\breve g$ is evaluated at $g$-invariant $\bk$ ($g\circ\bk=\bk$). For the reader's convenience, we present some commonly-used matrix representations for exactly degenerate bands in Tab.\ \ref{table:sewing-matrix}. The last column presents codimensions of the degeneracy manifold. For exactly-degenerate orbits in class 2, 4 and 5, $\A$ satisfies $\text{det}(\A){=}1$,[topofermi] with the assumption that spin-orbit coupling can be turned off continuously without changing the degeneracy of the orbit. 
 \label{table:codimension}}
\end{table*}

Generalizing our rotational-symmetric case studies in
Sec.\ \ref{sec:singleparameterrashba}-\ref{sec:rotsymmbreakdown}, we now present the symmetry-constrained codimensions for all ten[topoferm] symmetry classes of orbits. Depending on the class, the codimension is either $0,1,2$ or $3$. We begin by briefly reviewing the ten-fold symmetry classification of orbits in \s{sec:reviewtenfold}. We then study in Sec. \ref{sec:codimquasideg} five of the classes which is relevant to quasidegenerate orbits. The remaining five classes are relevant to exactly degenerate orbits, which we study in Sec. \ref{sec:codimexactdeg}.

@@

For these five classes (labelled 1, 3, 6, 7 and 10 according to the first column of Tab.\ \ref{table:codimension}), we list their corresponding \textit{symmetry-constrained codimensions} in the right-most column of Tab.\ \ref{table:codimension}; these numbers are derived in  App.\ \ref{app:codimension}. To clarify, we define the $g$-constrained codimension as the minimal codimension -- among all degeneracy manifolds -- as enforced by $g$ symmetry. To motivate `minimal' in this definition, we have found that two-fold rotational symmetry reduces the codimension of the $\A{=}{+}I$ manifold (to unity), but not the codimension of the $\A{=}{-}I$ manifold [cf.\ \q{proprashba}]; the rotation-constrained codimension is then unity. Generally, it is only in class 3 and 4 that different  degeneracy manifolds may have different codimensions. Each entry in the right-most column of Tab.\ \ref{table:codimension} equals the $g$-constrained codimension for any $g$ in that class (and possibly satisfying an additional constraint on the representation of $g$, as specified by the second-to-last column).

Of the five classes only three (1, 3 and 6) potentially have symmetry-constrained codimensions that are reduced from three: these three have in common that  $|g\circ\frako_0|{=}|\frako_0|$, and therefore the propagator $\A$ [cf.\ \q{eq:prop}] is self-constrained. To clarify, a self-constraint relates  the propagator $\A$ [cf.\ \q{eq:prop}] to itself or to its inverse $\A^{-1}$, depending respectively on whether $u{=}0$ or $1$.

@@

Class 3 corresponds to spatial symmetries that preserve the arrow of time and the orientation of $\frako_0$, as exemplified by rotational symmetry with rotational axis parallel to the field. The symmetry-constrained codimension here is unity, as has been modelled in Sec.\ \ref{sec:singleparameterrashba}-\ref{sec:rotsymmbreakdown}.

@@

\begin{table}
\begin{tabular*}{\columnwidth}{c@{\extracolsep{\fill}}ccc}
\hlineB{2.0}
class No. & symmetry  & origin of degeneracy & $\breve{g}\propto$ \\
\hline 
\multirow{2}{*}{1} & \multirow{2}{*}{$\mathfrak{r}_z$, $\mathfrak{r}_{z,\boldsymbol{a(b)}/2}$} & $T\mathfrak{i}$, spin SU(2) & $\sigma_z$ \\
\cline{3-4}
 & & $T\mathfrak{c}_{2z,\boldsymbol{c}/2}$ & $I$ \\
\hline
\multirow{3}{*}{5} & $\mathfrak{r}_{x(y),\boldsymbol{c}/2}$ & $T\mathfrak{i}$, spin SU(2), $T\mathfrak{c}_{2z,\boldsymbol{c}/2}$ & $\sigma_z$ \\
\cline{2-4}
& $\mathfrak{r}_{x(y)}$,$\mathfrak{r}_{x(y),\boldsymbol{b(a)}/2}$, & $T\mathfrak{i}$, spin SU(2) & $\sigma_z$ \\
\cline{3-4}
& $\mathfrak{c}_{2x(y)}$, $\mathfrak{c}_{2x(y),\boldsymbol{b(a)}/2}$ & $T\mathfrak{c}_{2z,\boldsymbol{c}/2}$ & $I$\\
\hlineB{2.0}
\end{tabular*}
\caption{Matrix representations of reflections (e.g., $\mathfrak{r}_x$) and rotations (e.g. $\mathfrak{c}_{2y}$) followed by possible half a lattice translations in $x,~y,~z$ directions (labeled by $\boldsymbol{a,~b,~c}$ respectively). This table is relevant for magnetic field in $\pm z$ direction. The two-fold exact degeneracy are assumed to be protected by either spatial-time inversion symmetry ($T\mathfrak{i}$), composition of time reversal and nonsymmorphic two fold rotation symmetry ($T\mathfrak{c}_{2z,\boldsymbol{c}/2}$) at $k_z=\pi$ or spin SU(2) (i.e., negligible SOC). \label{table:sewing-matrix}}
\end{table}

>>>>

\documentclass[aps, prb, showpacs, twocolumn, notitlepage, superscriptaddress, floatfix]{revtex4-1}
\usepackage{amsmath}
\usepackage{amssymb,amscd,bm,dsfont,wasysym,mathrsfs,latexsym,psfrag,accents,mathtools}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{color}
\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{boldline}
\usepackage{microtype}
\usepackage[normalem]{ulem}
\input{preamble}

@@

\begin{abstract}
\textbf{[abstract needs revamp]}
 In the semiclassical theory of Landau levels, a Bloch electron is driven by a magnetic field  along constant-energy band contours (in quasimomentum space) known as orbits. We introduce a quantization rule for Landau levels of nearly degenerate orbits which are split by spin-orbit coupling. The average area $S$ of two spin-orbit-split orbits is assumed much larger than the difference in areas ($\delta S$), and also much larger than the inverse square of the magnetic length ($\lmt$). By treating both $\delta S/S$ and $1/l^2S$ as small parameters with finite ratio $l^2\delta S$, our perspective diverges from and generalizes previous semiclassical theories.  $l^2\delta S$ is the dimensionless measure of the field-induced hybridization between spin-orbit-split orbits; we allow for $l^2\delta S$ to be comparable with the Zeeman coupling. In the opposing regimes $l^2\delta S {{\ll}} 1$  and $l^2\delta S{\gg}1$, we recover the known Onsager-Lifshitz-Roth rules for nondegenerate and degenerate orbits, respectively.  In the absence of crystallographic symmetry, three real parameters are needed to tune a spin-degeneracy between two Landau levels. Here, we identify symmetry classes for which such degeneracies can be tuned by varying a single parameter, which might be the magnitude or orientation of the field, or a bias voltage in tunneling spectroscopy.  We demonstrate the utility of our rule for the Rashba 2DEG -- with an arbitrarily oriented magnetic field, and with the possible addition of the Dresselhaus spin-orbit interaction.
\end{abstract}
\date{\today}

%, and the angular magnetoresistance oscillations of semiconductor bilayer heterostructures.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{RZ.png}
\caption{ Characterization of the Rashba 2DEG with an in-plane Zeeman field. (a) illustrates the trefoille-shaped cyclotron orbit at the energy of the II-Dirac point. The orbit is drawn over the  spin-splitting field ($d_x,d_y$) of the degeneracy-lifting Hamiltonian $\delta\var$ [cf.\ \q{eq:RZ-Hamiltonian}]. The wavefunction along the inner (resp.\ outer) component of the trefoille orbit has its pseudospin aligned parallel (resp.\ anti-parallel) to the spin-splitting field. The upper panel of (b) schematically illustrates  Landau levels near the II-Dirac point (in the weak-hybridization regime). The solid line represents the Onsager-Lifshitz quantization of the trefoille \textbf{[please change trefoille everywhere to trefoil]} orbit [cf.\ Eq.\ ({trefoillerule})], and the dashed line incorporates the first-order correction [cf.\ Eq.\ (\ref{RZfirstorderE})] due to tunneling. 
Lower panel of (b): red solid line shows the energetic difference between Landau levels determined by numerical diagonalization and Landau levels obtained by the trefoil quantization rule  [cf.\ Eq.\ ({trefoillerule})]; this solid line may be compared to the perturbative formula $\delta E_n$ [cf. Eq. (\ref{RZfirstorderE})], which is plotted as a black dashed line.  This comparison is made (at each $1/B$) for the Landau level closest to 0.45eV and with even Landau-level index $n$; the II-Dirac point is located at 0.5eV.  Parameters for lower panel of (b): $m{=}0.076m_0$, $\alpha{=} 1.5(10^{5}$cm/s$\cdot\hbar$), $Z_\parallel {=}$1meV. Out-of-plane Zeeman coupling is negligibly weak for the choice of $m$. (c) A sketch of cyclotron orbits of Rashba Hamiltonian with in-plane magnetic field. Inset: an enlarged view of near II-Dirac point. (d) For various configurations ($\alpha$, $Z_\parallel$) of $H_{RZ}$, we plot in left panel $\text{error}:=|\epsilon_\text{rule}-\epsilon_\text{exact}|/\var_c$ for a single Landau level that is closest to $\epsilon_0$ in Eq. (\ref{whereisdiracpoint}), where $\epsilon_\text{rule}$ is determined by Eq. (\ref{eq:rule}) and $\epsilon_{\text{exact}}$ is obtained by numerical diagonalization of the Peierls-Onsager Hamiltonian $H_{RZ}(\bK)$. Corresponding $l^2 \delta S^2/S$ are plotted in right panel. We only plotted in region where $0.4{<}\epsilon_0{<}1.0$(eV).\cite{RZfignote}, Units: $\alpha$($1.5\times 10^{5}$cm/s$\cdot\hbar$). Parameters for (d): $m=0.076m_0$, $l=100$\AA. \label{fig:RZ}}.
\end{figure}

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{blochsphere.png}
\caption{Solid-angle representation of  the single-band Berry phase $\phi_{\pm}^B$ for inner (subscript $-$, upper panel) and outer ($+$, lower panel) orbit, at various energies close to the Dirac point $\epsilon_0$ [cf.\ Eq. (\ref{whereisdiracpoint})]. The decomposition of the effective Hamiltonian [cf.\ Eq. (\ref{calHRZ})] into  Pauli matrices ($\calh{=}\sum_j\calh_j\sigma_j$) defines a three-vector $\boldsymbol{\calh}{=}({-}\alpha k_y$,$\alpha k_x{+}Z_\parallel,{-}Z_\perp$). $\phi_{\pm}^B$ equals half the solid angle subtended by   ${\mp} \boldsymbol{\calh}(\bk)$ [solid blue line] as $\bk$ is varied along the outer/inner orbit;[Berry] the orientation of the solid angle is indicated by blue dotted lines. The out-of-plane Zeeman interaction lifts $\boldsymbol{\calh}$ out of the x-y plane, but this occurs at most over a fraction ${\sim}(\bar{\mu}_{\sma{\text{min}}}m_0/m{\ll}1)$ of the cyclotron period. \label{fig:blochsphere}}
\end{figure}

@@

 To generalize the above discussion, let us parametrize ${\cal \bar{A}}$ by $d$ real coordinates  $\{r_1,\ldots,r_{d}\}{:}{=}\br $;  ${\cal \bar{A}}(\br){=}{\pm}I$ defines a set of submanifolds in $\br$-space which we refer to as degeneracy manifolds; in the above example, these manifolds comprise two points on $S^3$. The codimension of the degeneracy manifold  is defined as $d$ minus the dimension of said manifold; \textbf{[there are inconsistencies in these definitions, if we impose $d$ to be number of parameters (without imposing linear independence), then $d=4$ for canonical parametrization, and our definition of codimension as d minus dimension of degeneracy manifold gives 4. however, if we impose linearly-independent parameters, then we also run into trouble if we allow for symmetric coordinates that are not well-defined at points.]}\red{[First of all, I am not quite sure why the word linear independence is used in the current context. I would have used `independent', which means the parameters can be tuned simultaneously. One thing that you would probably dislike is that codimension defined by me would be dependent on the parameters $\br$, so it's OK that codimension in canonical parametrizaiton is three while codimension in exponential parametrization is one. The precise codimension would be the number of independent parameters needed to find a degeneracy. Although codimension depends on the parameters, the codimension found by us is generally universal. For example, for rotation symmetry in class 3, we parametrize $\A_{1/2}$ by canonical parametrization $\br_c$. There exists a mapping from physical parameters $\br$ to $\br_c$. Notice this mapping is not possible for exponential parametrization $\br_e$ since the exponential parametrization of identity is not unique. For a nonpathological parameter choice $\br$, the map is well behaved. The definition of `well-behaved': any $n\le 3$ dimensional manifold in $\br$ is mapped to $n$ dimensional manifold in $\br_c$. With this assumption, codimension of $\A$ in $\br_c$ and $\br$ are actually the same. Exponential parametrization is one example of pathological mapping, which maps to sphere to a point (identity). However, pathological parameter choice usually does not increase codimension. This can be argued in the following way: If the map from $\br$ to $\br_c$ maps a $n\le 3$ dimensional manifold in $\br$ to a $m<n$ dimensional manifold, we remove the corresponding manifold in $\br$. After all such manifolds removed, the map become what I defined as `well-behaved' and the codimension should be the same. If the manifold removed in $\br$ were mapped to $\A\propto I$, then the codimension decreases, which is the case in exponential parametrization. In the presentation, let's be vague and just say codimension is number of independent parameters needed to find a degeneracy.]} alternatively stated, the codimension is the number of linearly-independent parameters needed to fine-tune a degeneracy. Here and henceforth, we adopt the shorthand that  \textit{parameters} will always refer to real parameters, and \textit{codimension} always refers to a codimension of a degeneracy manifold.

@@

At energies close to the Dirac point ($E{-}\epsilon_0 {\ll} wm_2(2\epsilon_0/m_1)^{\sma{1/2}}$), there are two concentric orbits which (nearly) touch at both Dirac points, as illustrated in Fig.\ \ref{fig:Cn-breakdown}(a). For sufficiently weak field, interband tunneling is localized near either II-Dirac point, and occurs with the Landau-Zener probability:
\e{\rho^2=e^{-2\pi\bar{\mu}}, \; \bar{\mu}=\f{\big(Z_{\perp}^2+(m_1/m_2)^2(E-\epsilon_0)^2\big)l^2}{2w\sqrt{2E/m_1}}. \label{landauzenerc2}}
Equivalently, $\bar{\mu}{=}S_{\sma{\square}}l^2/8$, with $S_{\sma{\square}}$ the area of rectangle inscribed between the two hyperbolic arms [cf.\  Fig\ref{fig:Cn-breakdown}(a)]. Elsewhere, an electron undergoes adiabatic (band-conserving) dynamics. Due to $\mathfrak{c}_2$ symmetry, the time-evolution propagator over the full cyclotron period is simply the square of the propagator for half the period: 
\e{\A=\left(\matrixtwo{\tau e^{i\bar{\varphi}}}{-\rho}{\rho}{\tau e^{-i\bar{\varphi}}}{\diagmatrix{e^{i\tilde{\lambda}_-}}{e^{i\tilde{\lambda}_+}}} \right)^2. \label{propC2breakdown}}
The half-period propagator may be taken from \qq{scattmat}{propinplanezeeman}, with  $\bar{\mu}$ here given by \q{landauzenerc2}, and  $\tilde{\lambda}_{\pm}{=}{\pm}l^2\delta S/4$ plus a field-independent geometric phase.

@@

\begin{table*}
\begin{tabular*}{\textwidth}{c@{\extracolsep{\fill}}ccccccc}
\hlineB{2.0}
             class number & action on orbit & action on $\bk$ & $u$ & $s$ & quasidegenerate & $g$ & codimension\\
\hline
\multirow{2}{*}{1} & \multirow{10}{*}{$|g\circ \frako_0| = |\frako_0|$} & \multirow{4}{*}{$\forall \bk \in \frako_0$, $g\circ \bk=\bk$} & \multirow{2}{*}{0} & \multirow{2}{*}{0} & \multirow{2}{*}{$\surd$} & $\breve{g}\propto I$ & 3  \\
\cline{7-8}
 & & & & & & $\breve{g}\not\propto I$ & 1 \\
 \cline{4-8}
\multirow{2}{*}{2} & & & \multirow{2}{*}{0} & \multirow{2}{*}{1} & & $g^2=1$ & 1  \\
\cline{7-8}
 & & &  &  & & $g^2=-1$ & 3  \\
\cline{3-8}
3 & & \multirow{6}{*}{$\exists \bk \in \frako_0, g\circ \bk\ne\bk$} & 0 & 0 & $\surd$ & - & 1   \\
\cline{4-8}
4 & & & 0 & 1 & & - & 1   \\
\cline{4-8}
\multirow{2}{*}{5} & & & \multirow{2}{*}{1} & \multirow{2}{*}{0} & & $\breve{g} \propto \sigma_z$ & 2  \\
\cline{7-8}
 & & & &  & & $\breve{g} \not\propto \sigma_z$ & 0  \\
\cline{4-8}
\multirow{2}{*}{6} & & & \multirow{2}{*}{1} & \multirow{2}{*}{1} & \multirow{2}{*}{$\surd$} & $g^2=1$ & 2  \\
\cline{7-8}
& & & & & & $g^2=-1$ & 0  \\
\hline
7 & \multirow{4}{*}{$|g\circ \frako_0| \ne |\frako_0|$} & \multirow{4}{*}{-} & 0 & 0 & $\surd$ & - & 3 \\
\cline{4-8}
8 &  &  & 0 & 1 & & - & 3 \\
\cline{4-8}
9 &  &  & 1 & 0 & & - & 3 \\
\cline{4-8}
10 &  &  & 1 & 1 & $\surd$ & - & 3 \\
\hlineB{2.0}
\end{tabular*}
\caption{Symmetry-constrained codimensions for the two-by-two propagator $\A$. The first column numbers the symmetry class of $\A$. The second and third column defines how the symmetry acts on the cyclotron orbit. The next two columns  inform us if the mapping $\bk{\rightarrow}g{\circ}\bk$ [cf.\ \q{gmapk}]  preserves ($u{=}0$) or inverts ($u{=}1$) the orientation of the orbit, and if $g$ preserves ($s{=}0$) or inverts ($s{=}1$) the arrow of time. The sixth column indicates whether this  class is potentially a symmetry of quasidegenerate orbits. The seventh column specifies additional conditions (on the representation of $g$) which must be imposed to uniquely determine the codimension. $\breve{g}$ is defined as the diagonalized, unitary matrix representation of $g$ in the degenerate subspace of $\hat{H}_0$. In class 5, $\breve g$ is evaluated at $g$-invariant $\bk$ ($g\circ\bk=\bk$). For the reader's convenience, we present some commonly-encountered matrix representations for exactly-degenerate orbits in Tab.\ \ref{table:sewing-matrix}. The last column presents codimensions of the degeneracy manifold \textbf{[remind reader what is degeneracy manifold]}. For exactly-degenerate orbits in class 2, 4 and 5, $\A$ satisfies $\text{det}(\A){=}1$,[topofermi] with the assumption that spin-orbit coupling can be turned off continuously without changing the degeneracy of the orbit. 
 \label{table:codimension}}
\end{table*}

Generalizing our rotational-symmetric case studies in
Sec. \ref{sec:singleparameterrashba}-\ref{sec:rotsymmbreakdown}, we now present the symmetry-constrained codimensions for all ten[topoferm] symmetry classes of orbits. Depending on the class, the codimension is either $0,1,2$ or $3$. We begin by briefly reviewing the ten-fold symmetry classification of orbits in \s{sec:reviewtenfold}. We then study in Sec. \ref{sec:codimquasideg} five of the classes which is relevant to quasidegenerate orbits. The remaining five classes are relevant to exactly degenerate orbits, which we study in Sec. \ref{sec:codimexactdeg}.

@@

For these five classes (labelled 1, 3, 6, 7 and 10 according to the first column of Tab.\ \ref{table:codimension}), we list their corresponding \textit{symmetry-constrained codimensions} in the right-most column of Tab.\ \ref{table:codimension}; these numbers are derived in  App.\ \ref{app:codimension}. To clarify, we define the $g$-constrained codimension as a number which is robust to any perturbation of the Hamiltonian that preserves $g$ symmetry and the homotopy class of $\frako_0$. Each entry in the right-most column of Tab.\ \ref{table:codimension} equals the $g$-constrained codimension for any $g$ specified by the corresponding row.

%in that class (and possibly satisfying an additional constraint on the representation of $g$, as specified by the second-to-last column).

Of the five classes only classes 1, 3 and 6 potentially have symmetry-constrained codimensions that are reduced from three: classes 1, 3 and 6 have in common that  $|g\circ\frako_0|{=}|\frako_0|$, and therefore the propagator $\A$ [cf.\ \q{eq:prop}] is self-constrained. To clarify, a self-constraint relates  the propagator $\A$ [cf.\ \q{eq:prop}] to itself or to its inverse $\A^{-1}$, depending respectively on whether $u{=}0$ or $1$.

@@

Class 3 corresponds to spatial symmetries that preserve the arrow of time and the orientation of $\frako_0$, as exemplified by rotational symmetry with rotational axis parallel to the field. The symmetry-constrained codimension here is unity, as has been modelled in Sec.\ \ref{sec:singleparameterrashba}-\ref{sec:rotsymmbreakdown}.  Only in class 3 (among 1,3,6) do there exist different  degeneracy manifolds with different codimensions; in such case we define the symmetry-constrained codimension to be the minimal codimension among all manifolds. For example, we have found that two-fold rotational symmetry reduces the codimension of the $\A{=}{+}I$ manifold (to unity), but not the codimension of the $\A{=}{-}I$ manifold [cf.\ \q{proprashba}]. The rotation-constrained codimension is then unity, and hence the entry $1$ in the right-most column for class 3.

@@

\begin{table}
\begin{tabular*}{\columnwidth}{c@{\extracolsep{\fill}}ccc}
\hlineB{2.0}
class No. & symmetry  & origin of degeneracy & $\breve{g}\propto$ \\
\hline 
\multirow{2}{*}{1} & \multirow{2}{*}{$\mathfrak{r}_z$, $\mathfrak{r}_{z,\boldsymbol{a(b)}/2}$} & $T\mathfrak{i}$, spin SU(2) & $\sigma_z$ \\
\cline{3-4}
 & & $T\mathfrak{c}_{2z,\boldsymbol{c}/2}$ & $I$ \\
\hline
\multirow{3}{*}{5} & $\mathfrak{r}_{x(y),\boldsymbol{c}/2}$ & $T\mathfrak{i}$, spin SU(2), $T\mathfrak{c}_{2z,\boldsymbol{c}/2}$ & $\sigma_z$ \\
\cline{2-4}
& $\mathfrak{r}_{x(y)}$,$\mathfrak{r}_{x(y),\boldsymbol{b(a)}/2}$, & $T\mathfrak{i}$, spin SU(2) & $\sigma_z$ \\
\cline{3-4}
& $\mathfrak{c}_{2x(y)}$, $\mathfrak{c}_{2x(y),\boldsymbol{b(a)}/2}$ & $T\mathfrak{c}_{2z,\boldsymbol{c}/2}$ & $I$\\
\hlineB{2.0}
\end{tabular*}
\caption{Matrix representations of reflections (e.g., $\mathfrak{r}_x$), rotations (e.g. $\mathfrak{c}_{2y}$), glide reflections (e.g., $\mathfrak{r}_{x,\ba/2}$) and screw rotations (e.g., $\mathfrak{c}_{2y,\ba/2}$). The latter two nonsymmorphic elements involve  half a lattice translation in the $x,~y,$ or $z$ direction (labeled by $\boldsymbol{a,~b,~c}$ respectively). The $z$ axis of our Cartesian coordinate system is aligned parallel to the  magnetic field. This table is applicable to exactly-degenerate orbits whose degeneracy is protected by either spacetime inversion symmetry ($T\mathfrak{i}$), the composition of time reversal and nonsymmorphic two fold rotation symmetry ($T\mathfrak{c}_{2z,\boldsymbol{c}/2}$) [for orbits confined to the $k_z=\pi$ plane] or spin SU(2) symmetry (i.e., negligible spin-orbit coupling). \label{table:sewing-matrix}}
\end{table}
